|

### 2021-04-12 documentation/paper.Rmd
Combined results and constituencies files
Full recompilation

---
title: |
  | PollBasePro: Daily Estimates of Aggregate Voting Intention in Britain from `r format(min(britpol::pollbasepro$date), "%Y")` to the Present Day\thanks{This article benefitted from much useful feedback. We owe particular thanks to Chris Hanretty, who inspired this project and provided much useful feedback. We would also like to thank Hannah Willis, Y, and Z for offering their time to review our manuscript. Finally, we would like to extend a special thanks to all of the volunteers who work to maintain such a comprehensive list of British opinion polling data on Wikipedia.}
author: |
  | Jack Bailey\thanks{Research Associate, Department of Politics, University of Manchester, UK. If you have any comments or questions, feel free to contact me either by email (\href{mailto:jack.bailey@manchester.ac.uk}{jack.bailey@manchester.ac.uk}) or on Twitter (\href{https://www.twitter.com/PoliSciJack}{@PoliSciJack}).}, Mark Pack\thanks{President, The Liberal Democrats, UK}, & Luke Mansillo\thanks{PhD Candidate, Department of Government \& International Relations, University of Sydney, Australia}
date: |
  | \small This version compiled `r stringr::str_remove(format(Sys.time(), '%d %B %Y'), "^0")` using PollBasePro `r paste0("v", packageVersion("britpol"))`
abstract: |
  | Political scientists often use public opinion polls to test their theories. Yet these data present some difficulties. First, they are noisy. Second, they occur at irregular intervals. Third, they measure both public preferences and pollsters' survey design choices. We introduce a new dataset, PollBasePro, that accounts for these issues. It includes voting intention estimates for each of Britain's three largest parties on each day between the 1955 general election and the present. Given that the dataset covers `r format(nrow(britpol::pollbasepro), big.mark = ",")` days in total, it is more comprehensive than any existing time series of public opinion in Britain. As a result, we expect PollBasePro to be an important resource for students of British politics in the years to come.
  |
  | \textbf{\textsf{Keywords:}} Opinion polls; public opinion; British politics; Bayesian methods.
indent: yes
fontsize: 12pt
geometry: margin = 1.15in
subparagraph: yes
compact-title: false
bibliography: _assets/master.bib
biblio-style: _assets/apsr.bst
classoption: a4paper
linkcolor: black
urlcolor: violet
citecolor: black
output: 
  bookdown::pdf_document2: 
    latex_engine: xelatex
    toc: false
    keep_tex: false
    includes:
      in_header:
        - _assets/rmd-preamble.tex
    number_sections: false
    fig_caption: true
---

@@

Beleaguered politicians often remark that the "only poll that matters is the election itself". Like much that politicians say, this is not true. Though elections decide who governs, public opinion polls are important too: they let us track how popular parties are in the time between elections. Polling is especially important in Britain. Unlike in many countries, British governments have some control over scheduling elections. As a result, the governing party's standing is often one of the only factors that decides when it will hold an election [@smith2003].

@@

In this paper, we introduce a new dataset---*PollBasePro*---that overcomes these issues. It includes `r format(nrow(pollbasepro), big.mark = ",")` daily estimates of aggregate voting intention for each of Britain's three largest parties on each day between `r format(min(pollbasepro$date), "%d %B %Y")` and `r format(max(pollbasepro$date), "%d %B %Y")`. Covering `r format(nrow(pollbasepro), big.mark = ",")` days in total, PollBasePro permits a degree of specificity and flexibility beyond that of any existing time series data. In the sections that follow, we elaborate on the dataset. First, we describe the methods and the data that we use to derive our estimates. Next, we consider what PollBasePro reveals about British politics since 1955 and show how to use it to test hypotheses. We then provide some initial conclusions and remark on how we might develop PollBasePro in the future.

@@

Polling data have long informed political science research. For example, @goodhart1970 used Gallup and NOP polls to show that changes in the economy affect support for the incumbent party, thereby launching economic voting research. Though plentiful, these data come with some problems. As we mention above, three are most important.

First, polling data are noisy. This is an unavoidable consequence of sampling methodologies. As we cannot poll every person in a country, polls cannot provide exact estimates of public opinion. Rather, they represent draws from a distribution of *possible* estimates of public opinion. This noise is a type of measurement error, and failing to account for it can have important consequences. Where we use polls as an outcome, it reduces our statistical power. Where we use them as a predictor, it pulls real and existing effects towards zero.

Second, polling data occur at irregular intervals. What's more, they can cover any period of time from a single day to several weeks. Political scientists often assume that these figures measure public opinion on the poll's last day in the field. Of course, this is almost never true. It can also be a problem when major events occur partway through the data collection process. This is a major practical problem: polls can be difficult to match to other time series.

Third, polls do not only measure changes in aggregate preferences. They also include systematic biases due to the design choices of the companies that run them. In the past, these biases have been so large that they have cast doubt on the efficacy of the polling industry. At the 2015 UK general election, most polls suggested that Labour had a good chance of becoming the largest party. But, on the night, the Conservatives won a small majority instead [@sturgis2018; @prosser2018c, @mellon2017a]. These biases were even more serious at the 2019 Australian Federal Election. As in the British case, Labor had a healthy lead in the polls. Yet, again, the Coalition beat them by around 8 percentage points [@mansillo2020].

Our intention is simple: to produce voting intention estimates that account for these issues. To do so, we adapt the method in @jackman2005. Jackman's model has estimates for a given party start and end at known results from any given pair of elections. It then treats the party's level of support as a random walk between these two points. On any given day, the party's support depends on its support the day before, pollster-specific "house" effects, and random shocks. Jackman and Mansillo [-@mansillo2020; -@jackman2016] have used the model to track voting intention in Australia and @louwerse2016 have used it to estimate aggregate voting intention in Ireland. The full model specification is as follows^[Note that we provide a more in-depth explanation of how the model works and the steps that we have taken to produce our estimates in the user guide that accompanies the data.]:

@@

ggplot() +
  geom_density_2d(
    data = con %>% mutate(facet = "Conservative"),
    mapping = aes(x = vote, y = con_est),
    colour = pty_cols[1],
    alpha = .4
  ) +
  stat_smooth(
    data = con %>% mutate(facet = "Conservative"),
    mapping = aes(x = vote, y = con_est),
    colour = "black",
    fill = pty_cols[1],
    method = "lm",
    formula = y ~ x
  ) +
  # geom_text(
  #   data = con %>% mutate(facet = "Conservative"),
  #   mapping = 
  #     aes(
  #       x = .6,
  #       y = .05,
  #       label = 
  #         paste0(
  #           britpol:::in_text(pluck(posterior_samples(cor_con, "rescor"), 1)*100, suffix = "%", inside = F),
  #           "\n",
  #           "MAE = ", round(mae(cor_con$data$con_est, cor_con$data$vote), 2),
  #           ", RMSE = ", round(rmse(cor_con$data$con_est, cor_con$data$vote), 2)
  #           )
  #       ),
  #   hjust = 1,
  #   size = 2,
  #   family = "Cabin Regular"
  # ) +
  geom_density_2d(
    data = lab %>% mutate(facet = "Labour"),
    mapping = aes(x = vote, y = lab_est),
    colour = pty_cols[2],
    alpha = .4
  ) +
  stat_smooth(
    data = lab %>% mutate(facet = "Labour"),
    mapping = aes(x = vote, y = lab_est),
    colour = "black",
    fill = pty_cols[2],
    method = "lm",
    formula = y ~ x
  ) +
  # geom_text(
  #   data = lab %>% mutate(facet = "Labour"),
  #   mapping = 
  #     aes(
  #       x = .6,
  #       y = .05,
  #       label = 
  #         paste0(
  #           britpol:::in_text(pluck(posterior_samples(cor_lab, "rescor"), 1)*100, suffix = "%", inside = F),
  #           "\n",
  #           "MAE = ", round(mae(cor_lab$data$lab_est, cor_lab$data$vote), 2),
  #           ", RMSE = ", round(rmse(cor_lab$data$lab_est, cor_lab$data$vote), 2)
  #           )
  #     ),
  #   hjust = 1,
  #   size = 2,
  #   family = "Cabin Regular"
  # ) +
  geom_density_2d(
    data = lib %>% mutate(facet = "Liberal (Various Forms)"),
    mapping = aes(x = vote, y = lib_est),
    colour = pty_cols[3],
    alpha = .4
  ) +
  stat_smooth(
    data = lib %>% mutate(facet = "Liberal (Various Forms)"),
    mapping = aes(x = vote, y = lib_est),
    colour = "black",
    fill = pty_cols[3],
    method = "lm",
    formula = y ~ x
  ) +
  # geom_text(
  #   data = lib %>% mutate(facet = "Liberal (Various Forms)"),
  #   mapping = 
  #     aes(
  #       x = .6,
  #       y = .05,
  #       label = 
  #         paste0(
  #           britpol:::in_text(pluck(posterior_samples(cor_lib, "rescor"), 1)*100, suffix = "%", inside = F),
  #           "\n",
  #           "MAE = ", round(mae(cor_lib$data$lib_est, cor_lib$data$vote), 2),
  #           ", RMSE = ", round(rmse(cor_lib$data$lib_est, cor_lib$data$vote), 2)
  #           )
  #     ),
  #   hjust = 1,
  #   size = 2,
  #   family = "Cabin Regular"
  # ) +
  facet_wrap(~ facet) +
  scale_y_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_x_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  coord_cartesian(
    ylim = c(0, 0.62),
    xlim = c(0, 0.62)
  ) +
  labs(x = "Jennings and Wlezien's 'Timeline of Elections' Data", y = "PollBasePro Estimates") +
  theme_bailey()
```

We use the Wikipedia data to cover the period from 2010 onwards. This is a pragmatic choice. Volunteers update the data in real time and include sample sizes that are missing from PollBase. This is important as our model assumes that we know the sampling error present in each estimate. For the PollBase data, we impute likely sample sizes to allow us to estimate these errors^[Again, see the user guide that accompanies the data for more information on this process.]. But as the Wikipedia data include sample sizes, imputation is not necessary. While some might doubt Wikipedia's reliability, we do not think that it is an issue. Polling figures are verifiable and likely of interest only to a very small group of people. Wikipedia's coverage of public opinion polling since 2010 is also of a very high quality, and almost all figures on the website including links to external data sources that corroborate them.

We validate them against Jennings and Wlezien's "Timeline of Elections" dataset [-@jennings2016a]. These data contain `r format(nrow(timeline[timeline$party == 1, ]), big.mark = ",")` polls from Britain from `r str_remove(format(min(as_date(timeline$date)), "%d %B %Y"), "^0")` to `r str_remove(format(max(as_date(timeline$date)), "%d %B %Y"), "^0")`. Given that our data are so comprehensive, it is likely that most polls appear in both datasets. Still, the Timeline data provide a good test as Jennings and Wlezien compiled them independently. As figure \@ref(fig:val-plot) shows, our estimates are well validated. Correlations between the two series are strong and positive. Their mean absolute error (MAE) and root-mean-square error (RMSE) are also low in all cases. The Conservatives showed a correlation of `r britpol:::in_text(cor_mods$tl_con*100, suffix = "%", inside = F)`, an MAE of `r paste0(round(cor_mods$mae_con, 2), " percentage points")`, and an RMSE of `r round(cor_mods$rmse_con, 2)`; Labour, a correlation of `r britpol:::in_text(cor_mods$tl_lab*100, suffix = "%", inside = F)`, an MAE of `r paste0(round(cor_mods$mae_lab, 2), " points")`, and an RMSE of `r round(cor_mods$rmse_lab, 2)`; and the Liberals, a correlation of `r britpol:::in_text(cor_mods$tl_lib*100, suffix = "%", inside = F)`, an MAE of `r paste0(round(cor_mods$mae_lib, 2), " points")`, and an RMSE of `r round(cor_mods$rmse_lib, 2)`.

@@

Figure \@ref(fig:time-plot) shows that our estimates track British political history very well. Between 1955 and 1980, we observe the heyday of the two-party system. Here, around `r scales::percent(median(pollbasepro$lab_est[pollbasepro$date %in% seq.Date(as_date("1955-01-01"), as_date("1980-01-01"), "days")]), accuracy = .1)` of the electorate supported Labour and another `r scales::percent(median(pollbasepro$con_est[pollbasepro$date %in% seq.Date(as_date("1955-01-01"), as_date("1980-01-01"), "days")]), accuracy = .1)` the Conservatives. In the 1980s, we see the rise and fall of the SDP-Liberal alliance. Labour’s slow rise to power in 1997 soon follows, as does their loss of support over the next decade and a half. More recently, the data show a brief “blip” in Liberal support to coincide with “Cleggmania” in 2010, Labour’s surprise surge in 2017, and the Conservative's landslide support in 2019.

@@

\newpage
\begin{landscape}
```{r time-plot, fig.cap = paste("PollBasePro includes", format(nrow(pollbasepro), big.mark = ","), "daily estimates of aggregate voting intention for each of Britain's largest parties: the Conservatives, Labour, and the Liberals in their various forms."), fig.width = 10, fig.height = 6, fig.align = "centre", echo = F}

@@

kable(
  vitals,
  align = "lrrrrrr",
  format = "latex",
  label = "vitals-tab",
  booktabs = TRUE,
  escape = FALSE,
  caption = paste0("Overall summary of voting intention estimates, ", format(min(pollbasepro$date), "%Y"), " to ", format(max(pollbasepro$date), "%Y"))
  ) %>% 
  kable_styling(position = "center")

@@

@@

According to the literature, there are good reasons why the relationship between the two variables might be positive and good reasons why it might be negative. The literature on "rally 'round the flag" effects would favour a *positive* relationship. Here, voters rush to support the governing party in times of crisis, such as wars [@kuijpers2019; @mueller1970]. The literature on retrospective voting [@healy2013; @fiorina1981], instead, would favour a *negative* relationship. It holds that voters reward governments for positive policy outcomes and punish them for negative ones.

@@

It is clear then that PollBasePro has allowed us to reveal an important---if surprising---answer to the question that we posed above: despite Britain having one of the world's highest levels of COVID-related deaths per capita [@ritchie2021], these untimely deaths appear to have *benefited* the fortunes of the governing Conservative Party. Still, the relationship has changed over time and, a year into the pandemic, has reversed direction. We might, thus, expect any further deaths attributable to COVID-19 to erode the Conservative's ample base of support. Yet given the pace of Britain's vaccination programme [@ukgovernment2021a], it is possible that the Conservatives have reaped the rewards without facing any punishment.

@@

In this paper, we introduce PollBasePro: the most comprehensive time series of British voting intention data assembled to date. The dataset contains `r format(nrow(britpol::pollbasepro), big.mark = ",")` daily voting intention estimates for each of Britain's three largest parties, beginning at the 1955 General Election on `r format(min(pollbasepro$date), "%d %B %Y")` and ending on `r str_remove(format(max(pollbasepro$date), "%d %B %Y"), "^0")`. Furthermore, our data are well validated and follow the course of British political history. We therefore expect PollBasePro to become a valuable resource for students of British politics in the years to come.

@@

@@

>>>>

---
title: |
  | PollBasePro: Daily Estimates of Aggregate Voting Intention in Britain from `r format(min(britpol::pollbasepro$date), "%Y")` to the Present Day\thanks{This article benefitted from much useful feedback. We owe particular thanks Chris Hanretty, who inspired this project and provided much useful feedback. We would also like to thank Hannah Willis, Chris Wlezien, and Patrick English for offering their time to review our manuscript. Finally, we would like to extend a special thanks to all of the volunteers who work to maintain such a comprehensive list of British opinion polling data on Wikipedia.}
author: |
  | Jack Bailey\thanks{Research Associate, Department of Politics, University of Manchester, UK. If you have any comments or questions, feel free to contact me either by email (\href{mailto:jack.bailey@manchester.ac.uk}{jack.bailey@manchester.ac.uk}) or on Twitter (\href{https://www.twitter.com/PoliSciJack}{@PoliSciJack}).}, Mark Pack\thanks{President, The Liberal Democrats, UK}, & Luke Mansillo\thanks{PhD Candidate, Department of Government \& International Relations, University of Sydney, Australia}
date: |
  | \small This version compiled `r stringr::str_remove(format(Sys.time(), '%d %B %Y'), "^0")` using PollBasePro `r paste0("v", packageVersion("britpol"))`
abstract: |
  | Political scientists often use public opinion polls to test their theories. Yet these data present some difficulties. First, they are noisy. Second, they occur at irregular intervals. Third, they measure both public preferences and pollsters' survey design choices. We introduce a new dataset, PollBasePro, that accounts for these issues. It includes voting intention estimates for each of Britain's three largest parties on each day between the 1955 general election and the present. As the dataset covers `r format(nrow(britpol::pollbasepro), big.mark = ",")` days in total, it is perhaps more comprehensive than any other existing time series of British political opinion. We then use our estimates to test a question of pressing importance: how daily deaths attributable to COVID-19 have influenced support for the governing Conservative Party.
  |
  | \textbf{\textsf{Keywords:}} Opinion polls; public opinion; British politics; Bayesian methods.
indent: yes
fontsize: 12pt
geometry: margin = 1.15in
subparagraph: yes
compact-title: false
bibliography: _assets/master.bib
biblio-style: _assets/apsr.bst
classoption: a4paper
linkcolor: black
urlcolor: violet
citecolor: black
output: 
  bookdown::pdf_document2: 
    latex_engine: xelatex
    toc: false
    keep_tex: false
    includes:
      in_header:
        - _assets/rmd-preamble.tex
    number_sections: false
    fig_caption: true
---

@@

Beleaguered politicians often remark that the "only poll that matters is the election itself". Like much that politicians say, this is not true. Though elections decide who governs, public opinion polls are important too: they tell us how popular parties are in the time between elections. Polling is especially important in Britain. Unlike in some countries, British governments can control the election schedule. As a result, the governing party's standing is often one of the only factors that decides when an election will occur [@smith2003].

@@

We introduce a new dataset---*PollBasePro*---that overcomes these issues. It includes daily estimates of aggregate voting intention for each of Britain's three largest parties on each day between `r format(min(pollbasepro$date), "%d %B %Y")` and `r format(max(pollbasepro$date), "%d %B %Y")`. Covering `r format(nrow(pollbasepro), big.mark = ",")` days in total, PollBasePro permits a degree of specificity and flexibility beyond that of any existing time series. In the sections that follow, we elaborate on the dataset. First, we describe the methods and the data that we use to derive our estimates. Next, we consider what PollBasePro reveals about British politics since 1955 and show how to use it to test answer a question of pressing concern: how do deaths related to COVID-19 affect public support for the governing Conservative Party? We show that COVID-related deaths have led to a "rally 'round the flag" effect that has boosted Conservative support, though that this relationship has weakened, then reversed, as the pandemic has progressed. Finally, we provide some initial conclusions and then remark on how we might develop PollBasePro over the coming years.

@@

Polls have long informed political science. @goodhart1970, for instance, used Gallup and NOP polls to show that economic change affects incumbent party support, launching the economic vote. Yet though plentiful, these data have three important problems.

First, polling data are noisy. This is an unavoidable consequence of sampling methodologies. As we cannot poll every person in a country, polls cannot provide exact estimates of public opinion. Rather, they represent draws from a distribution of *possible* estimates of public opinion. Once introduced into formal statistical models, this sampling error becomes a form of measurement error. And failing to account for it can have serious consequences. Where we use polls as an outcome, it reduces our statistical power. Where we use them as a predictor, it pulls real and existing effects towards zero.

Second, polling data occur at irregular intervals, covering any period of time from a single day to several weeks. Political scientists often assume that these figures measure public opinion on the poll's last day in the field. Of course, that is almost never true. It can also be a problem when major events occur partway through the data collection process. This is a major practical problem: polls can be difficult to match to other time series.

Third, polls do not only measure changes in aggregate political preferences. They also measure systematic biases that arise due to the design choices of the companies that run them. In the past, these biases have been so large that they have cast doubt on the efficacy of the polling industry. At the 2015 UK general election, most polls suggested that Labour had a good chance of becoming the largest party. But, on the night, the Conservatives won a small majority instead [@sturgis2018; @prosser2018b; @mellon2017]. These biases were even more serious at the 2019 Australian Federal Election. As in the British case, Labor had a healthy lead in the polls. Yet, again, the conservative Coalition beat them by around 8 percentage points [@mansillo2020].

Our intention is simple: to produce voting intention estimates that account for these issues. To do so, we adapt the method in @jackman2005. Jackman's model has estimates for a given party start and end at known results from any given pair of elections. It then treats the party's level of support as a random walk between these two points. On any given day, the party's support depends on its support the day before, pollster-specific "house" effects, and random shocks. Jackman and Mansillo [-@mansillo2020; -@jackman2016] have used the model to track voting intention in Australia and @louwerse2016 have used it to estimate aggregate voting intention in Ireland. The full model specification is as follows, though note that we also provide a more complete explanation of how the model works and the steps that we have taken to produce our estimates in the accompanying appendix:

@@

ggplot() +
  geom_density_2d(
    data = con %>% mutate(facet = "Conservative"),
    mapping = aes(x = vote, y = con_est),
    colour = pty_cols[1],
    alpha = .7
  ) +
  stat_smooth(
    data = con %>% mutate(facet = "Conservative"),
    mapping = aes(x = vote, y = con_est),
    colour = "black",
    fill = pty_cols[1],
    method = "lm",
    formula = y ~ x
  ) +
  # geom_text(
  #   data = con %>% mutate(facet = "Conservative"),
  #   mapping = 
  #     aes(
  #       x = .6,
  #       y = .05,
  #       label = 
  #         paste0(
  #           britpol:::in_text(pluck(posterior_samples(cor_con, "rescor"), 1)*100, suffix = "%", inside = F),
  #           "\n",
  #           "MAE = ", round(mae(cor_con$data$con_est, cor_con$data$vote), 2),
  #           ", RMSE = ", round(rmse(cor_con$data$con_est, cor_con$data$vote), 2)
  #           )
  #       ),
  #   hjust = 1,
  #   size = 2,
  #   family = "Cabin Regular"
  # ) +
  geom_density_2d(
    data = lab %>% mutate(facet = "Labour"),
    mapping = aes(x = vote, y = lab_est),
    colour = pty_cols[2],
    alpha = .4
  ) +
  stat_smooth(
    data = lab %>% mutate(facet = "Labour"),
    mapping = aes(x = vote, y = lab_est),
    colour = "black",
    fill = pty_cols[2],
    method = "lm",
    formula = y ~ x
  ) +
  # geom_text(
  #   data = lab %>% mutate(facet = "Labour"),
  #   mapping = 
  #     aes(
  #       x = .6,
  #       y = .05,
  #       label = 
  #         paste0(
  #           britpol:::in_text(pluck(posterior_samples(cor_lab, "rescor"), 1)*100, suffix = "%", inside = F),
  #           "\n",
  #           "MAE = ", round(mae(cor_lab$data$lab_est, cor_lab$data$vote), 2),
  #           ", RMSE = ", round(rmse(cor_lab$data$lab_est, cor_lab$data$vote), 2)
  #           )
  #     ),
  #   hjust = 1,
  #   size = 2,
  #   family = "Cabin Regular"
  # ) +
  geom_density_2d(
    data = lib %>% mutate(facet = "Liberal (Various Forms)"),
    mapping = aes(x = vote, y = lib_est),
    colour = pty_cols[3],
    alpha = .4
  ) +
  stat_smooth(
    data = lib %>% mutate(facet = "Liberal (Various Forms)"),
    mapping = aes(x = vote, y = lib_est),
    colour = "black",
    fill = pty_cols[3],
    method = "lm",
    formula = y ~ x
  ) +
  # geom_text(
  #   data = lib %>% mutate(facet = "Liberal (Various Forms)"),
  #   mapping = 
  #     aes(
  #       x = .6,
  #       y = .05,
  #       label = 
  #         paste0(
  #           britpol:::in_text(pluck(posterior_samples(cor_lib, "rescor"), 1)*100, suffix = "%", inside = F),
  #           "\n",
  #           "MAE = ", round(mae(cor_lib$data$lib_est, cor_lib$data$vote), 2),
  #           ", RMSE = ", round(rmse(cor_lib$data$lib_est, cor_lib$data$vote), 2)
  #           )
  #     ),
  #   hjust = 1,
  #   size = 2,
  #   family = "Cabin Regular"
  # ) +
  facet_wrap(~ facet) +
  scale_y_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_x_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  coord_cartesian(
    ylim = c(0, 0.62),
    xlim = c(0, 0.62)
  ) +
  labs(x = "Jennings and Wlezien's 'Timeline of Elections' Data", y = "PollBasePro Estimates") +
  theme_bailey()
```

We use the Wikipedia data to cover the period from 2010 onwards. This is a pragmatic choice. Volunteers update the data in real time and include sample sizes that are missing from PollBase. This is important as our model assumes that we know the sampling error present in each estimate. For the PollBase data, we impute likely sample sizes to allow us to estimate these errors^[Again, see the appendix for more information on this process.]. But as the Wikipedia data include sample sizes, imputation is not necessary. While some might doubt Wikipedia's reliability, we do not think that it is an issue. Polling figures are verifiable and likely of interest only to a very small group of people. Wikipedia's coverage of public opinion polling since 2010 is also of a very high quality, and almost all figures on the website including links to external data sources that corroborate them.

We validate them against Jennings and Wlezien's "Timeline of Elections" dataset [-@jennings2016a]. These data contain `r format(nrow(timeline[timeline$party == 1, ]), big.mark = ",")` polls from Britain from `r str_remove(format(min(as_date(timeline$date)), "%d %B %Y"), "^0")` to `r str_remove(format(max(as_date(timeline$date)), "%d %B %Y"), "^0")`. Given that our data are so comprehensive, it is likely that most polls appear in both datasets. Even so, the Timeline data provide a good test as Jennings and Wlezien compiled them independently. As figure \@ref(fig:val-plot) shows, our estimates are well validated. Correlations between the two series are strong and positive. Their mean absolute error (MAE) and root-mean-square error (RMSE) are also low in all cases. The Conservatives showed a correlation of `r britpol:::in_text(cor_mods$tl_con*100, suffix = "%", inside = F)`, an MAE of `r paste0(round(cor_mods$mae_con, 2), " percentage points")`, and an RMSE of `r round(cor_mods$rmse_con, 2)`; Labour, a correlation of `r britpol:::in_text(cor_mods$tl_lab*100, suffix = "%", inside = F)`, an MAE of `r paste0(round(cor_mods$mae_lab, 2), " points")`, and an RMSE of `r round(cor_mods$rmse_lab, 2)`; and the Liberals, a correlation of `r britpol:::in_text(cor_mods$tl_lib*100, suffix = "%", inside = F)`, an MAE of `r paste0(round(cor_mods$mae_lib, 2), " points")`, and an RMSE of `r round(cor_mods$rmse_lib, 2)`.

@@

Figure \@ref(fig:time-plot) shows that our estimates track British political history well. From 1955 to 1980, we see the heyday of the two-party system. Here, around `r scales::percent(median(pollbasepro$lab_est[pollbasepro$date %in% seq.Date(as_date("1955-01-01"), as_date("1980-01-01"), "days")]), accuracy = 1)` of the electorate supported Labour and another `r scales::percent(median(pollbasepro$con_est[pollbasepro$date %in% seq.Date(as_date("1955-01-01"), as_date("1980-01-01"), "days")]), accuracy = 1)` the Conservatives. In the 1980s, we see the rise and fall of the SDP-Liberal alliance. Labour’s slow rise to power in 1997 soon follows, as does their loss of support over the next decade and a half. More recently, the data show a “blip” in Liberal support that coincides with “Cleggmania” in 2010, Labour’s surge in 2017, and the Conservative landslide in 2019.

@@

@@

kable(
  vitals,
  align = "lrrrrrr",
  format = "latex",
  label = "vitals-tab",
  booktabs = TRUE,
  escape = FALSE,
  caption = paste0("Overall summary of daily voting intention estimates, ", format(min(pollbasepro$date), "%Y"), " to ", format(max(pollbasepro$date), "%Y"))
  ) %>% 
  kable_styling(position = "center")

@@

\begin{landscape}
```{r time-plot, fig.cap = paste("PollBasePro includes", format(nrow(pollbasepro), big.mark = ","), "daily estimates of aggregate voting intention for each of Britain's largest parties: the Conservatives, Labour, and the Liberals in their various forms."), fig.width = 10, fig.height = 6, fig.align = "centre", echo = F}

@@

It is possible to propagate this error forward because PollBasePro includes *two* estimates for each party for each day that the data covers: estimates of public support for each party *and* an estimate of the standard error of this estimate. Table \@ref(tab:head-tab) shows the first 5 rows of the PollBasePro dataset. Variables ending "_est" show estimated voting intention and those ending "_err" their associated error. Notice that the error on the first date is zero. This is because this is the date of the 1955 general election, so we have absolute certainty what the result really was. As the days pass, this error gradually increases, reflecting our own epistemic uncertainty. These errors provide useful information in their own right, so it is prudent to also include them in our analysis.

```{r head-table, echo = F}

# Get table head

pbp_head <- 
  head(
    pollbasepro,
    n = 5
    ) %>% 
  mutate_if(
    is.numeric,
    function(x) round(x, 3)
  )

# Add latex code to variable names

names(pbp_head) <- paste0("\\textsf{\\textbf{" ,names(pbp_head), "}}")
names(pbp_head) <- str_replace(names(pbp_head), "_", "\\\\_")

kable(
  pbp_head,
  align = "ccccccc",
  format = "latex",
  label = "head-tab",
  booktabs = TRUE,
  escape = FALSE,
  caption = paste0("The first 5 rows of the PollBasePro dataset")
  ) %>% 
  kable_styling(position = "center")

According to the literature, there are good reasons why the relationship between the two variables might be positive and good reasons why it might be negative. The literature on "rally 'round the flag" effects would favour a *positive* relationship. Here, voters rush to support the governing party in times of crisis, such as wars [@kuijpers2019; @mueller1970]. The literature on retrospective voting [@healy2013; @fiorina1981], instead, would favour a *negative* relationship. It holds that voters reward governments for positive policy outcomes and punish them for negative ones (i.e. for avoidable deaths).

@@

It is clear then that PollBasePro has allowed us to reveal an important---if surprising---answer to the question that we posed above: despite Britain having one of the world's highest levels of COVID-related deaths per capita [@ritchie2021], these untimely deaths appear to have *benefited* the fortunes of the governing Conservative Party. Yet this relationship has changed over time and, a year into the pandemic, has reversed direction. We might, thus, expect any further deaths attributable to COVID-19 to erode the Conservative's ample base of support. Yet given the pace of Britain's vaccination programme [@ukgovernment2021a], it is possible that the Conservatives have reaped the electoral rewards of the crisis without risk of punishment.

@@

In this paper, we introduce PollBasePro: the most comprehensive time series of British voting intention data assembled to date. The dataset contains `r format(nrow(britpol::pollbasepro), big.mark = ",")` daily voting intention estimates for each of Britain's three largest parties, beginning at the 1955 General Election on `r format(min(pollbasepro$date), "%d %B %Y")` and ending on `r str_remove(format(max(pollbasepro$date), "%d %B %Y"), "^0")`. Furthermore, our data are well validated and follow the course of British political history.

@@

No matter how PollBasePro develops in the future, we are clear that it provides many opportunities in the present. We have discussed many of the academic opportunities that the data provide above. But our estimates might stand to improve the quality of polling reporting too. The stories that we tell ourselves to explain the polls often come to shape our politics [@barnfield2020]. And many of these stories emerge from the simple averages that many commentators use to make sense of the polls. As we mention above, our estimates account for shortcomings that these simple rolling averages do not. PollBasePro might, therefore, stand to benefit not only academic researchers, but also the quality of political debate in Britain too. As such, we expect PollBasePro to become a valuable resource for students of British politics in the years to come.

@@

# Appendix: Estimating Daily Voting Intention

We adapt Jackman's [-@jackman2005] method to derive our daily estimates. Still, there are issues specific to our case that we must first overcome. We elaborate on our choices below.

## Imputating Missing Sample Sizes

Our data do not include sample sizes before the 2010 general election. This is a problem, as our model requires that we know this information. To solve this problem, we use data from Jennings and Wlezien's [-@jennings2016a] "Timeline of Elections" dataset. Though less comprehensive than PollBase, these data do include information on sample sizes. What's more, they also include data from countries other than Britain. This lets us pool all available information to improve our estimates.

Sample sizes are count data. As such, we use the following multilevel Poisson regression model to impute likely sample sizes for all of our pre-2010 polling data:

\begin{align*}
n_{i} &\sim \mathrm{Poisson}(\lambda_{i}) \srlab{Likelihood function} \\
log(\lambda_{i}) &= \alpha_{Country[i]} + \beta_{Country[i]} T_{i} \srlab{Linear model on $\lambda$} \\
\begin{bmatrix} \alpha_{Country} \\ \beta_{Country} \end{bmatrix} &\sim \mathrm{MVNormal}(\begin{bmatrix} \alpha \\ \beta \end{bmatrix}, \textbf{S}) \srlab{Multivariate prior on varying effects} \\
\textbf{S} &= \begin{pmatrix} \sigma_{\alpha} & 0 \\ 0 & \sigma_{\beta} \end{pmatrix} \textbf{R} \begin{pmatrix} \sigma_{\alpha} & 0 \\ 0 & \sigma_{\beta} \end{pmatrix} \srlab{Covariance matrix on varying effects} \\
\alpha &\sim \mathrm{Normal}(7, 0.5) \srlab{Prior on average intercept, $\alpha$} \\
\beta &\sim \mathrm{Normal}(0, 0.1) \srlab{Prior on average slope, $\beta$} \\
\sigma_{\alpha} &\sim \mathrm{Exponential}(10) \srlab{Prior on uncertainty in the intercepts, $\sigma_{\alpha}$} \\
\sigma_{\beta} &\sim \mathrm{Exponential}(10) \srlab{Prior on uncertainty in the slopes, $\sigma_{\beta}$} \\
\textbf{R} &\sim \mathrm{LKJ}(2) \srlab{Prior on correlation matrix, $\textbf{\textrm{R}}$} \\
\end{align*}

```{r n-plot, fig.cap = paste0("Imputed sample sizes in Britain between ", format(min(samplesizes$date), "%Y"), " and ", format(max(samplesizes$date), "%Y"), ", estimated using sample size data from Jennings and Wlezien's \"Timeline of Elections\" dataset (2016)."), fig.width = 6, fig.height = 3, fig.align = "centre", echo = F}

# Create sample size plot
